---
title: "Lab1_inClass"
author: "Rae Fuhrman"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:

- basic review (data wranginling)
- intro to spatial data in R with sf
- intro to shiny apps

1. Attach packages

```{r load packages}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)

```

2. Review reading, exploring, wrangling, and visualizing data using data on wine quality and prices.

a. read in data
```{r}

wine <- read_csv("wine_data.csv") #don't use read.csv, it's depracated
#exploratory, so use console rather than View so it doesn't come up everytime I knit

```

b. wrangling - don't overwrite my original data as best practice
```{r}

wine_new <- wine %>% 
  select(country, province, winery, region_1, points, price) %>% #select columns, also used for ordering
  rename(state = province) %>%  #single equal sets something, == looks for a match
  filter(state == "California" | state == "Oregon" | state == "Washington") %>%  #| means OR
  mutate(ppd = points/price) %>%  #ppd as new column
  arrange(-ppd) #- means arrange by decreasing

```

c. use group_by() and summarize() to find summary statistics by group
```{r}

wine_summary <- wine_new %>% 
  group_by(state) %>%  #r is now thinking to calculate by this state column, this dataset currently looks exactly the same but it will recognize each state as a unique level  
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE)
  )
```

d. a few gg graphics
```{r}

wine_hists <- ggplot(wine_new, aes(x = ppd))+#piping is adding additional functions, + in ggplot means we are adding additional levels
  geom_histogram(aes(fill=state))+#use fill in histogram to distinguish color
  facet_wrap(~state, scales = "free")#sets the scale appropriately for the scale of each rather than all the same
wine_hists

```


new graph: scatterplot graph of price vs. points, with color determined by state, adding a linear model trendline on top of it
```{r}

wine_points <- ggplot(wine_new, aes(x = price, y = points))+
  geom_point(aes(color = state), alpha=0.5)+ #points use color, bars use fill as fill color and color as line color, if referencing a variable use aes if not don't use it
  geom_smooth(color="black", method = "lm", se = FALSE)+
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange"))+#c stands for combination
  facet_wrap(~state)+
  theme_light()
wine_points

```

2. California counties (spatial data) using sf

```{r}

ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016") #"." means we are already there in working directory
#plot(ca_counties)
ca_land <- ca_counties %>% 
  select(NAME, ALAND) #didn't ask it to keep the geometry but sf knows you probably want to keep it

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  rename(NAME = COUNTY)#rename county column to be name so the datasets can be merged

ca_df <- full_join(ca_land, ca_pop_inc) %>%  #full join keeps every row even if it doesn't have a match in the other data frame
  select(NAME, MedFamilyIncome)
```

D. Use geom_sf to make a map in ggplot
```{r}

ca_income <- ggplot(ca_df)+
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% 
  addTiles() %>%  #need to change the CRS to match what leaflet uses which is CRS 4326
  addPolygons()
  
```





